---
title: "Advanced Usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6,
  fig.height=4
)
```

# Advanced SuperPlots

In this vignette, we will explore some of the more advanced features of
SuperPlotR.

But first, we need to deal with how to simply scale the plot so it looks good in
your paper!

## Sizing the SuperPlot

The default setting is to use point sizes of 2 for the data and 3 for the
summary points. The font size is 12. This looks good in the RStudio viewer, but
is not well suited to a figure which is likely to be very small.

```{r sizing}
library(SuperPlotR)
# the default plot
superplot(lord_jcb, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)")
# the same plot but with custom sizing
superplot(lord_jcb, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)",
          size = c(0.8,1.5), fsize = 9)
```

This does not look great in the viewer, but it will look better in a figure.

```{r sizing2, fig.height=1.8, fig.width=1.8}
library(ggplot2)
# the same plot but with custom sizing
superplot(lord_jcb, "Speed", "Treatment", "Replicate",
               ylab = "Speed (µm/min)", size = c(0.8,1.5), fsize = 9)
ggsave("plot.pdf", width = 88, height = 50, units = "mm") # final size
```

This is preferable to

```{r sizing3, fig.height=1.8, fig.width=1.8}
library(ggplot2)
# the same plot but with custom sizing
superplot(lord_jcb, "Speed", "Treatment", "Replicate",
               ylab = "Speed (µm/min)")
ggsave("plot.pdf", width = 88, height = 50, units = "mm") # final size
```

## Customising the SuperPlot

A couple of simple tweaks: an x label can be added, and the transparency of
points can be altered like this.

```{r tweaks}
superplot(lord_jcb, "Speed", "Treatment", "Replicate",
          xlab = "Drug", ylab = "Speed (µm/min)", alpha = c(0.3,1))
```

SuperPlotR returns a ggplot object which can be customised how you like. For
example, the theme can be overridden like this:

```{r theme}
p <- superplot(lord_jcb, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)")
p + theme_minimal()
```

It can also accept a ggplot object using the `gg` parameter, and then add a
SuperPlot to it (within reason!).
For example, you might want to plot something behind the SuperPlot.

```{r ggplot}
p <- ggplot() +
  geom_hline(yintercept = 20, linetype = "dashed", col = "grey")
superplot(lord_jcb, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)", gg = p)
```

## Ordering the x-axis

This is best done by reordering the levels of the factor in the input dataframe
before calling `superplot`.

```{r order}
df <- lord_jcb
df$Treatment <- factor(df$Treatment, levels = c("Drug", "Control"))
superplot(df, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)")
```

It is also possible to reorder the Replicates using a similar strategy. You
might want to do this so that the order of colours and shapes matches a
different order to the default. Another way to achieve the same thing is to
supply a reordered colour palette to `superplot`.
