---
title: "Advanced Usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6,
  fig.height=4
)
```

# Going deeper: advanced SuperPlots

In this vignette, we will explore some of the more advanced features of
SuperPlotR.

But first, we need to deal with how to simply scale the plot so it looks good in
your paper!

## Sizing the SuperPlot

The default setting is to use point sizes of 2 for the data and 3 for the
summary points. The font size is 12. This looks good in the RStudio viewer, but
is not well suited to a figure which is likely to be very small.

```{r sizing}
library(SuperPlotR)
# the default plot
superplot(lord_jcb, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)")
# the same plot but with custom sizing
superplot(lord_jcb, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)",
          size = c(0.8,1.5), fsize = 9)
```

This does not look great in the viewer, but it will look better in a figure.

```{r sizing2, fig.height=1.8, fig.width=1.8}
library(ggplot2)
# the same plot but with custom sizing
superplot(lord_jcb, "Speed", "Treatment", "Replicate",
               ylab = "Speed (µm/min)", size = c(0.8,1.5), fsize = 9)
ggsave("plot.pdf", width = 88, height = 50, units = "mm") # final size
```

This is preferable to

```{r sizing3, fig.height=1.8, fig.width=1.8}
library(ggplot2)
# the same plot but with custom sizing
superplot(lord_jcb, "Speed", "Treatment", "Replicate",
               ylab = "Speed (µm/min)")
ggsave("plot.pdf", width = 88, height = 50, units = "mm") # final size
```

## Customising the SuperPlot

A couple of simple tweaks: an x label can be added, and the transparency of
points can be altered like this.

```{r tweaks}
superplot(lord_jcb, "Speed", "Treatment", "Replicate",
          xlab = "Drug", ylab = "Speed (µm/min)", alpha = c(0.3,1))
```

SuperPlotR returns a ggplot object which can be customised how you like. For
example, the theme can be overridden like this:

```{r theme}
p <- superplot(lord_jcb, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)")
p + theme_minimal()
```

It can also accept a ggplot object using the `gg` parameter, and then add a
SuperPlot to it (within reason!).
For example, you might want to plot something behind the SuperPlot.

```{r ggplot}
p <- ggplot() +
  geom_hline(yintercept = 20, linetype = "dashed", col = "grey")
superplot(lord_jcb, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)", gg = p)
```

## Facetting

Facetting is partially supported. In the simple scenario of facetting by
replicate, you can do this:

```{r facet}
p <- superplot(lord_jcb, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)")
p + facet_wrap(~Replicate)
```

If we had another variable that we wished to facet by, it is possible, but with
a limitation: The summary points will remain as the replicate summary and
therefore will be the same for each facet. This may not be what you want.

```{r facet2}
df <- cbind(lord_jcb, other = rep(c("A", "B"), 150))
p <- superplot(df, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)")
p + facet_wrap(~other)
```

The best way to do this and have the correct summary points is to make two
superplots after filtering the data for A and B, and then combine them using
`{patchwork}` or similar package.

## Ordering the x-axis

This is best done by reordering the levels of the factor in the input dataframe
before calling `superplot`.

```{r order}
df <- lord_jcb
df$Treatment <- factor(df$Treatment, levels = c("Drug", "Control"))
superplot(df, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)")
```

## Conditions and Replicates are categories

If you have a dataset where the conditions and replicates are numeric,
SuperPlotR converts them to character for plotting. This is the intended
behaviour for a SuperPlot.

Currently, if your data has Conditions that are numeric, for example time points, and these
are unevenly spaced, e.g. 0, 1, 3, 9 hours; then these will appear evenly spaced
on the x-axis.

## Adding statistics

This is currently not supported due to the number of different ways that users
may wish to add them. However, it is possible to add them manually to the
SuperPlot object.
