---
title: "Advanced Usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6,
  fig.height=4
)
```

# Advanced SuperPlots

In this vignette, we will explore some of the more advanced features of
SuperPlotR.

But first, we need to deal with how to simply scale the plot so it looks good in
your paper!

## Sizing the SuperPlot

The default setting is to use point sizes of 2 for the data and 3 for the
summary points. The font size is 12. This looks good in the RStudio viewer, but
is not well suited to a figure which is likely to be very small.

```{r sizing}
library(SuperPlotR)
# the default plot
superplot(lord_jcb, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)")
# the same plot but with custom sizing
superplot(lord_jcb, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)",
          size = c(0.8,1.5), fsize = 9)
```

This does not look great in the viewer, but it will look better in a figure.

```{r sizing2, fig.height=1.8, fig.width=1.8}
library(ggplot2)
# the same plot but with custom sizing
superplot(lord_jcb, "Speed", "Treatment", "Replicate",
               ylab = "Speed (µm/min)", size = c(0.8,1.5), fsize = 9)
ggsave("plot.pdf", width = 88, height = 50, units = "mm") # final size
```

This is preferable to

```{r sizing3, fig.height=1.8, fig.width=1.8}
library(ggplot2)
# the same plot but with custom sizing
superplot(lord_jcb, "Speed", "Treatment", "Replicate",
               ylab = "Speed (µm/min)")
ggsave("plot.pdf", width = 88, height = 50, units = "mm") # final size
```

## Customising the SuperPlot

A couple of simple tweaks: an x label can be added, and the transparency of
points can be altered like this.

```{r tweaks}
superplot(lord_jcb, "Speed", "Treatment", "Replicate",
          xlab = "Drug", ylab = "Speed (µm/min)", alpha = c(0.3,1))
```

SuperPlotR returns a ggplot object which can be customised how you like. For
example, the theme can be overridden like this:

```{r theme}
p <- superplot(lord_jcb, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)")
p + theme_minimal()
```

It can also accept a ggplot object using the `gg` parameter, and then add a
SuperPlot to it (within reason!).
For example, you might want to plot something behind the SuperPlot.

```{r ggplot}
p <- ggplot() +
  geom_hline(yintercept = 20, linetype = "dashed", col = "grey")
superplot(lord_jcb, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)", gg = p)
```

## Ordering the x-axis

This is best done by reordering the levels of the factor in the input dataframe
before calling `superplot`.

```{r order}
df <- lord_jcb
df$Treatment <- factor(df$Treatment, levels = c("Drug", "Control"))
superplot(df, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)")
```

It is also possible to reorder the Replicates using a similar strategy. You
might want to do this so that the order of colours and shapes matches a
different order to the default. Another way to achieve the same thing is to
supply a reordered colour palette to `superplot`.

## Getting information about your SuperPlot

Having made your SuperPlot, you might want to know a bit more about it.
For example, you might wonder which replicate is which or perhaps you received
a warning that some replicates are missing some conditions.

You can set the option `info = TRUE` when you call `superplot` to get more
detailed information.

```{r info}
superplot(lord_jcb, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)",
          info = TRUE)
```

When this is set, the SuperPlot will contain a legend, otherwise, the legend is
not shown by default. This is because the legend is not very useful in most
cases, as the colours and shapes are already shown in the plot. However, if you
want to see the legend, you can either set `info = TRUE` or use the append one
the output using `+ theme(legend.position = "right")`.

### Retrieving the summary data

If you need to retrieve the summary data used to create the SuperPlot, you can
use the `get_sp_summary` function. This will return a data frame with the
summary data used to create the SuperPlot.

```{r summary}
summary_data <- get_sp_summary(lord_jcb, "Speed", "Treatment", "Replicate")
head(summary_data)
```

### Finding representative datapoints

If you want to find the representative datapoints for each condition, then you
can use the `representative()` function. This is handy if the data come from
a set of images and you'd like to show a representative image in the figure.

The function returns a data frame with the datapoints ranked by closeness to
the summary of each replicate or condition (see `?representative` for more
information). It also prints the top ranked datapoint to the console.

```{r representative}
representative_data <- representative(lord_jcb, "Speed", "Treatment", "Replicate")
head(representative_data)
```

In this example, the dataset has no label column, so the row number is used as
the label. If you have a label column, you can specify it using the `label`
parameter. The label could be a filename or other identifier to identify where
the datapoint came from.

```{r representative_label}
# Assuming lord_jcb has a column "FileName" with the labels
example <- lord_jcb
example$FileName <- paste0("Image_", seq_len(nrow(example)), ".tif")
representative_data <- representative(example, "Speed", "Treatment", "Replicate",
                                      label = "FileName")
```
