---
title: "Making non-SuperPlots - FlatPlots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Making non-SuperPlots - FlatPlots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Making non-SuperPlots - FlatPlots

SuperPlotR is designed to make SuperPlots, but sometimes you just want a simple
plot. This vignette will show you how to make a "FlatPlot" using SuperPlotR.

In contrast to SuperPlots, which emphasise the experimental replicates,
FlatPlots have a flat structure, where the replicates are not shown. They can
be used to look at data from a single experiment, or when the replicates are the
individual data points.

```{r flatplot}
library(SuperPlotR)
flatplot(lord_jcb, "Speed", "Treatment", ylab = "Speed (um/min)")
```

Many of the arguments are the same as for `superplot()`, but the `replicate`
argument is not used.

```{r flatplot2}
flatplot(lord_jcb, "Speed", "Treatment", ylab = "Speed (um/min)",
         colour = "rl_green")
```

The control of colour is by a single colour, which can be a hex code or one of
our lab's publication colour palette.

```{r flatplot3}
flatplot(lord_jcb, "Speed", "Treatment", ylab = "Speed (um/min)",
         colour = "rl_red", stats = TRUE)
```

We can request statistical testing as for SuperPlots, but the p-values will be
calculated for the whole dataset, not for each replicate.

```{r flatplot4}
flatplot(lord_jcb, "Speed", "Treatment", ylab = "Speed (um/min)",
         colour = "rl_red", size = 4, alpha = 0.25,
         bars = "mean_ci")
```

In this example, we have increased the size of the points, made them slightly
more transparent, and added error bars to show the mean and 95% confidence
interval.

