---
title: "Going Deeper"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Going Deeper}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6,
  fig.height=4
)
```

# Going deeper: advanced SuperPlots

In this vignette, we will explore some of the more advanced features of SuperPlotR.

But first, we need to deal with how to simply scale the plot so it looks good in
your paper!

## Sizing the SuperPlot

The default setting is to use point sizes of 2 for the data and 3 for the
summary points. The font size is 12. This looks good in the RStudio viewer, but
is not well suited to a figure which is likely to be very small.

```{r sizing}
library(SuperPlotR)
# the default plot
superplot(lord_jcb, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)")
# the same plot but with custom sizing
superplot(lord_jcb, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)",
          size = c(0.8,1.5), fsize = 9)
```

This does not look great in the viewer, but it will look better in a figure.

```{r sizing2, fig.height=1.8, fig.width=1.8}
library(ggplot2)
# the same plot but with custom sizing
superplot(lord_jcb, "Speed", "Treatment", "Replicate",
               ylab = "Speed (µm/min)", size = c(0.8,1.5), fsize = 9)
ggsave("plot.pdf", width = 88, height = 50, units = "mm")
```

## Customising the SuperPlot

A couple of simple tweaks: and x label can be added and transparency of points
can be altered like this.

```{r tweaks}
superplot(lord_jcb, "Speed", "Treatment", "Replicate",
          xlab = "Drug", ylab = "Speed (µm/min)", alpha = c(0.3,1))
```

SuperPlotR returns a ggplot object which can be customised how you like. For
example, the theme can be overridden like this:

```{r theme}
p <- superplot(lord_jcb, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)")
p + theme_minimal()
```

It can also accept a ggplot object and add the SuperPlot to it (within reason!).
For example, you might want to plot some things behind the data points.

```{r ggplot}
p <- ggplot() +
  geom_hline(yintercept = 20, linetype = "dashed", col = "grey")
superplot(lord_jcb, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)", gg = p)
```

## Facetting

Facetting is partially supported. In the simple scenario of facetting by
replicate, you can do this:

```{r facet}
p <- superplot(lord_jcb, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)")
p + facet_wrap(~Replicate)
```

If we had another variable that we wished to facet by, it is possible but with
a limitation. The summary points will be the same for each facet, which may not
be what you want.

```{r facet2}
df <- cbind(lord_jcb, other = rep(c("A", "B"), 150))
p <- superplot(df, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)")
p + facet_wrap(~other)
```

The best way to do this and have the correct summary points is to make two
superplots after filtering the data for A and B, and then combine them using
`patchwork` or similar package.

## Ordering the x-axis

This is best done by reordering the levels of the factor in the input dataframe
before calling `superplot`.

```{r order}
df <- lord_jcb
df$Treatment <- factor(df$Treatment, levels = c("Drug", "Control"))
superplot(df, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)")
```
