---
title: "Limitations to SuperPlotR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Limitations to SuperPlotR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6,
  fig.height=4
)
```

# Current limitations

There are some limitations to SuperPlotR. This vignette covers them and offers
solutions where possible.

## Facetting

Facetting is _partially supported_. In the simple scenario of facetting by
replicate, you can do this:

```{r facet}
library(SuperPlotR)
library(ggplot2)
p <- superplot(lord_jcb, "Speed", "Treatment", "Replicate",
               ylab = "Speed (µm/min)")
p + facet_wrap(~Replicate)
```

If we had another variable that we wished to facet by, then facetting by a this
variable is possible, but with a limitation: The summary points will remain as
the replicate summary and therefore will be the same for each facet.
This may not be what you want.

```{r facet2}
df <- cbind(lord_jcb, other = rep(c("A", "B"), 150))
p <- superplot(df, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)")
p + facet_wrap(~other)
```

**Solution:** The best way to do achieve "facetting" currently (and have the correct summary
points) is to make a SuperPlot for each variable, i.e. after filtering the data frame for A and B,
and then combine them using `{patchwork}` or similar package.

## Conditions and Replicates are categories

If you have a dataset where the conditions and replicates are numeric,
SuperPlotR converts them to character for plotting. This is the intended
behaviour for a SuperPlot.

Currently, if your data has Conditions that are numeric, for example time
points, and these are unevenly spaced, e.g. 0, 1, 3, 9 hours; then these will
appear evenly spaced on the x-axis. This is a limitation of the current
implementation. Raise an issue if you would like this to be changed.

## Adding statistics

This is currently not supported due to the number of different ways that users
may wish to add p-values or stars and the different tests that would need to be
accommodated. However, it is possible to add them manually to the SuperPlot
object. Again, raise an issue if you would like this to be changed.
