---
title: "Making non-SuperPlots - PiePlots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pieplot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(SuperPlotR)
```

# Making non-SuperPlots - FlatPlots

For convenience, SuperPlotR also provides a function to make pie charts. Called
`pieplot()`, it takes a vector of values and a vector of colours, and returns a
pie chart using `ggplot2`. It can also take a second vector of values to create
a pie chart with two layers.

```{r pieplot1}
pieplot(x1 = c(123, 456),
        cols = c("#44aa99", "#117733"))
```
```{r pieplot2}
pieplot(x1 = c(50 - 20, 20, 80, 180 - 80),
        cols = c("#bbbbbb", "#44aa99", "#117733", "#dddddd"),
        x2 = c(100, 130))
```

The `cols` argument should be a character vector of colours, which can be
hex codes or one of our lab's publication colour palette. The function will
convert the colours to a format that `ggplot2` can use.

```{r pieplot3}
pieplot(x1 = c(50 - 20, 20, 80, 180 - 80),
        cols = c("#bbbbbb", "rl_blue", "#117733", "lavender"),
        x2 = c(10, 13), label = "Example Pie Chart")
```

Note that the values for the second pie chart layer do not have the same
as the first layer. The function will automatically scale the second layer.
