---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6,
  fig.height=4
)
```

# A basic SuperPlot

Example data from the original paper is included in the package. We'll use this
for demonstration. The data is called `lord_jcb`.
A basic SuperPlot can be generated with the following code:

```{r setup}
library(SuperPlotR)
superplot(lord_jcb, "Speed", "Treatment", "Replicate", ylab = "Speed (µm/min)")
```

This will generate a SuperPlot of the `Speed` data, with `Treatment` on the
x-axis and `Replicate` is used to group the data. The y-axis label is set to
"Speed".

An equivalent call in long form is:

```{r setup2, echo = TRUE, results = "hide"}
superplot(
  df = lord_jcb,
  meas = "Speed",
  cond = "Treatment",
  repl = "Replicate",
  ylab = "Speed (µm/min)"
)
```

# Customising the SuperPlot

A separate vignete describes how to customise the sizing of the plot. Here, we
will look at major customisation options.

## Data distribution

The underlying data can be rendered as violins,

```{r violin}
superplot(lord_jcb, "Speed", "Treatment", "Replicate",
          ylab = "Speed (µm/min)", datadist = "violin")
```

or with a jittered scatter plot, the default (shown in the first example on this
page) uses `sina` which is a jittered scatter plot with a density estimate.

```{r jitter}
superplot(lord_jcb, "Speed", "Treatment", "Replicate",
          ylab = "Speed (µm/min)", datadist = "jitter")
```

## Colours

The colours of the points can be changed. The default is Paul Tol's bright
palette (`"tol_bright"`) but you can select from:

- `"tol_bright"`
- `"tol_vibrant"`
- `"tol_muted"`
- `"tol_light"`
- or Color Universal Design (`"cud"`)

```{r colours}
superplot(lord_jcb, "Speed", "Treatment", "Replicate",
          ylab = "Speed (µm/min)", pal = "cud")
```

It's possible to supply your own palette as a vector of colours (in hex or
description format)

```{r custom_palette}
superplot(lord_jcb, "Speed", "Treatment", "Replicate",
          ylab = "Speed (µm/min)", pal = c("#ff0000", "blue", "green"))
```

## Summary points

The mean of each replicate is shown by default. You can show the median instead

```{r median}
superplot(lord_jcb, "Speed", "Treatment", "Replicate",
          ylab = "Speed (µm/min)", rep_summary = "rep_median")
```

## Bars

The default SuperPlot does not include bars to indicate the mean ± error. They
can be added by using the `bars` argument. Options are:

- `"mean_sd"` = mean ± standard deviation
- `"mean_sem"` = mean ± standard error of the mean
- `"mean_ci"` = mean ± 95% confidence interval

Note that these bars show the mean ± error of the replicate means not the
underlying data points.

```{r bars}
superplot(lord_jcb, "Speed", "Treatment", "Replicate",
          ylab = "Speed (µm/min)", bars = "mean_sd")
```

## Data point shapes

By default, all data points are circles. If you want to use different shapes for
different conditions, you can use the `shapes` argument. Set to `TRUE` to use
different shapes for each replicate.

```{r shapes}
superplot(lord_jcb, "Speed", "Treatment", "Replicate",
          ylab = "Speed (µm/min)", shapes = TRUE)
```

If you'd like to link the summary points, set `linking = TRUE`.

```{r linking}
superplot(lord_jcb, "Speed", "Treatment", "Replicate",
          ylab = "Speed (µm/min)", linking = TRUE)
```

Linkages can be made when there are more than two groups. To demonstrate this
we will use some toy data:

```{r linking2}
set.seed(123)
example <- data.frame(meas = rep(rep(c(10, 9, 6, 6), each = 25), 4) + rnorm(400),
                      cond = rep(rep(c("ctrl", "drug 1", "drug 2", "drug 3"), each = 25), 4),
                      expt = rep(c("exp1","exp2","exp3","exp4"), each = 100))
superplot(example, "meas", "cond", "expt",
          linking = TRUE)
```

# Saving the SuperPlot

The SuperPlot can be saved as a PDF or PNG file using the `ggsave` function from
the `ggplot2` package.

```{r save}
library(ggplot2)
p <- superplot(lord_jcb, "Speed", "Treatment", "Replicate",
               ylab = "Speed (µm/min)")
ggsave("superplot.pdf", p)
```

# More customisation

A separate vignette describes how to customise the SuperPlot further.
See `vignette("advanced")`.

# Session Info

```{r}
sessionInfo()
```

